---
title: "Project 2 - 1.Pharma"
subtitle: "Ben Inbar"
output: html_document
date: "10/9/22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Choose any three of the “wide” datasets identified in the Week 6 Discussion items. (You may use your own dataset; please don’t use my Sample Post dataset, since that was used in your Week 6 assignment!) For each of the three chosen datasets: 1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You’re encouraged to use a “wide” structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below. 2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!] 3) Perform the analysis requested in the discussion item.

<div style="margin-top:50px;">
Read in Jawaid Hakim's pharmaceutical drug spending .csv and check.
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
data <- read.csv("https://raw.githubusercontent.com/beninbar/DATA-607/main/Project%202/Pharma%20spend%20by%20country.csv")
head(data)
```

<div style="margin-top:50px;">
Fortunately, this dataset is already relatively tidy and well-suited to analysis. But one problem that is immediately clear is that the number of observations for each country differs. Perhaps worse, some countries have data for differing years. 47 different years are represented, but not all countries are represented in each year.
```{r}
data |> summarize(countries = count(data, LOCATION))
data |> summarize(countries = count(data, TIME))
```

<div style="margin-top:50px;">
We can try plotting this anyway, per Jawaid's suggestions, just to see what it looks like visually. Note that figuring out how best to add labels took a while! They work, but in future would like cleaner ones. To position them properly, I generated a summarized tibble extracted from the larger dataset to pull from.
```{r}
for_label <- data |> group_by(LOCATION) |> summarize(year=max(TIME), totalspend=max(TOTAL_SPEND))

ggplot(data=data, aes(x=TIME, y=TOTAL_SPEND, group=LOCATION)) +
  geom_line() +
  geom_text(data=for_label, aes(x=year, y=totalspend + 4000, label=LOCATION))
```

<div style="margin-top:50px;">
The USA is clearly far and away the biggest phara drug spender. We can remove USA to get a better look at the trends in the rest of the countries.
```{r}
noUSA <- filter(data, LOCATION!="USA")
for_labelnoUSA <- noUSA |> group_by(LOCATION) |> summarize(year=max(TIME), totalspend=max(TOTAL_SPEND))

ggplot(data=noUSA, aes(x=TIME, y=TOTAL_SPEND, group=LOCATION)) +
  geom_line() +
  geom_text(data=for_labelnoUSA, aes(x=year, y=totalspend + 4000, label=LOCATION)) +
  ggtitle("NoUSA")
```

<div style="margin-top:50px;">
For Jawaid's other suggestion of comparing growth in spend between countries, I took the difference between the max and min in TOTAL_SPEND for each country across all years, and graphed.
```{r}
growth <- data |> group_by(LOCATION) |> summarize(growth=max(TOTAL_SPEND) - min(TOTAL_SPEND))
ggplot(data=growth, aes(x=LOCATION, y=growth)) +
  geom_bar(stat="identity") +
  geom_text(data=growth, aes(x=LOCATION, y=growth + 4000, label=LOCATION), size=3) +
  theme(axis.text.x=element_blank(), axis.title.x=element_blank()) +
  labs(y="Change in spending - all years")
```

##### Once again we see that the US far eclipses other countries, not only in total spend but in growth in spending over time.