---
title: "DATA 607 Final project"
author: "Ben Inbar, Josh Forster"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="margin-top:50px;">
##### In this project, we look at immigration data from various sources to determine how the US and each individual state compares for migrant employment.
##### First, we took data from OECD.org, which compiles data for OECD member countries.

<div style="margin-top:50px;">
First, we pulled OECD data [here](https://stats.oecd.org/Index.aspx?DataSetCode=MIG) and uploaded to [Github](https://github.com/beninbar/DATA-607/blob/main/Final%20project/OECD.zip).

```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(DescTools)
library(knitr)
 library(usmap)
library(gridExtra)

temp <- tempfile()
oecd <- download.file("https://github.com/beninbar/DATA-607/blob/main/Final%20project/data/OECD.zip?raw=TRUE", temp, mode = "wb")
oecd <- unzip(temp, "OECD.csv")
oecd <- read.csv(oecd)

kable(head(oecd))
nrow(oecd)
oecd |> group_by(Variable) |> summarize()
```
<div style="margin-top:50px;">
As we can see, the data is in long format. The "Variable" column is of particular interest, and for immigration specifically, the "Inflows of foreign population by nationality" subset.

Let's clean the dataframe: get rid of duplicate/unnecessary columns, subset to values of only inflows of foreign population, and aggregate inflows per country for all countries of birth (first left hand column).
```{r warning=FALSE, message=FALSE}
oecd <- oecd |> filter(Variable=="Inflows of foreign population by nationality") |> select(Country, Year, Value)

aggregated <- oecd |> group_by(Country, Year) |> summarize(Totals = sum(Value))
kable(head(aggregated))
```

<div style="margin-top:50px;">
Let's visualize this with a multi-line chart.
```{r}
library(ggrepel)

for_label <- aggregated |> filter(Year=="2015")

ggplot(data=aggregated, aes(x=Year, y=Totals, color=Country)) +
  geom_line() +
  ggtitle("Inflows of foreign population per country, 2000-2021") +
  theme(axis.title.y=element_blank(), legend.position="none", panel.grid.minor = element_blank()) +
  scale_y_continuous(labels = scales::comma, expand=c(0, 100000)) +
  scale_x_continuous(expand=c(.01, 0)) + 
  geom_label_repel(data = subset(for_label, Country==c("Germany", "United States")), aes(label=Country), nudge_x=2, nudge_y=400000,     segment.color=NA)
```

<div style="margin-top:50px;">
It's clear that the United States and Germany are two of the highest inflow countries by far. This tracks with the migrant crisis in 2015, when Germany opened its borders to unprecedented numbers of Middle Eastern and African refugees.

We can also run an F-test with Dunnett's post-hoc test to compare the United States to all other OECD countries. We're using a Dunnett's posthoc test since in this case we have many comparisons (35 countries compared to each other), but we're only interested in finding out if one of them (the United States) is significantly different.
```{r}
anova <- aov(Totals ~ Country, aggregated)
summary(anova)

set.seed(42)
DunnettTest(x=aggregated$Totals, g=aggregated$Country, control="United States")
```

<div style="margin-top:50px;">
#### Indeed, the only other country that shows less significant of a difference in migration (but notably, is still significant), is Germany!
<div style="margin-top:50px;">
  
#### US State Comparison
  
  #### Load Data
```{r load_data}
link <- 'https://raw.githubusercontent.com/beninbar/DATA-607/main/Final%20project/data/state_immigration_stats.csv'
state_data <- read_csv(link,show_col_types=FALSE)
```

#### How do wages for US-born vs foreign born differ based on available data

```{r}
median_wages <- state_data |>
    group_by(state) |>
    summarise(foreign_born_median_wages = mean(foreign_median_wages),
              us_born_median_wages = mean(us_median_wages),
              diff_wages = us_born_median_wages-foreign_born_median_wages)

ggplot(data=median_wages,mapping=aes(x=us_born_median_wages,y=foreign_born_median_wages))+
    geom_point() +
    geom_jitter() +
    geom_text(aes(label=state),color='black',size=4,vjust=1) +
    geom_smooth(method='lm') +
    labs(title= '2019 Median Annual Wages by Birth',subtitle = 'Source: Migration Policy Institute via American Community Survey')

cor(median_wages$us_born_median_wages,median_wages$foreign_born_median_wages)


t.test(median_wages$diff_wages,conf.level = .95)
```

From reviewing the scatterplot of median pages by birth, it is clear there is a very strong linear pattern between the two suggesting that wages are not that different in gross terms between US and foreign born residents. The correlation between the two variables is very high which might indicate and would lead us to hypothesize that state employment factors are a much stronger predictor of employment for all of its residents. After running a single T-test it is also clear that there is a statistically significant difference between the two median wages with a 95% confidence interval for the true population estimates at (3743.393,7318.019)

#### Top/Bottom 5 Wage Differential

```{r}
median_wages |>
    arrange(-diff_wages) |>
    head(5)

median_wages |>
    arrange(-diff_wages) |>
    tail(5)
```
While it is interesting to see this information in a list perhaps showing the full map visualization will be a better holistic view of the whole country's statistics.

#### US and Foreign Born Maps



#### Differential Map

```{r}
plot_usmap(
    data = median_wages, values = "diff_wages", color = "white"
  ) + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Median Wages Difference (2019)", label = scales::comma
  ) + 
  labs(title = "Median Wage Comparison", subtitle = "US - Foreign Wages.") +
  theme(legend.position = "right")
```

#### Separated Median Wage Map

```{r}
us_map <- plot_usmap(
    data = median_wages, values = "us_born_median_wages", color = "white"
  ) + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Median US Born Wages (2019)", label = scales::comma
  ) + 
  labs(title = "Median Wages (US Born)") +
  theme(legend.position = "right")
foreign_map <- plot_usmap(
    data = median_wages, values = "foreign_born_median_wages", color = "white"
  ) + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Median Foreign Born Wages (2019)", label = scales::comma
  ) + 
  labs(title = "Median Wages (Foreign Born)") +
  theme(legend.position = "right")

grid.arrange(us_map,foreign_map,ncol=2)
```

#### Data Cleanup
Prepare scraped data for better formattings and correct data types
```{r}
state_bin <- state_data |> filter(order!=0) |>
    mutate(earning_bin = str_trim(str_replace_all(header,c('Earned'='', ', or incurred a loss'=''))),
           foreign_born_perc = as.numeric(str_replace(foreign_born,'%',''))/100,
           us_born_perc = as.numeric(str_replace(us_born,'%',''))/100)
```


#### Foreign Workers in Highest Wage Band

```{r}
state_bin |> filter(earning_bin=='$75,000 or more') %>%
    ggplot(aes(reorder(state,foreign_born_perc),foreign_born_perc,fill='blue')) +
    theme(axis.text.x = element_text(angle=90,hjust=0.9),legend.position = 'none')+ 
    geom_bar(stat='identity') +
    labs(x='States',y='Foreign Born (%)', title='Percentage of Foreign Born making $75,000+')
    
```

#### Number of states in each binned wage range

```{r}
state_bin_top <- state_bin |> 
    group_by(state) |> 
    mutate(top_rank_foreign = rank(-foreign_born_perc,ties.method = 'last'),
           top_rank_us=rank(-us_born_perc)) |> 
    pivot_longer(cols=c('top_rank_foreign','top_rank_us'),names_to='measure_name',values_to='measure_values') |>
    filter(measure_values==1)    


most_common_bins <- state_bin |> 
    group_by(state) |> 
    mutate(top_rank_foreign = rank(-foreign_born_perc),
           top_rank_us=rank(-us_born_perc)) |> 
    pivot_longer(cols=c('top_rank_foreign','top_rank_us'),names_to='measure_name',values_to='measure_values') |>
    filter(measure_values==1) %>%
    
    group_by(earning_bin,measure_name) |>
    summarise(count = n()) %>%
    ggplot(aes(reorder(earning_bin,count),count,fill=measure_name)) +
    geom_bar(stat='identity',position='dodge') +
    theme(axis.text.x = element_text(angle=90,hjust=0.9))+
    labs(x='Earnings Bin',y='Number of States', title='Highest Frequency Wage Bins')
 
 most_common_bins + scale_fill_discrete(name="Category",labels=c('Most Common Foreign Born','Most Common US Born'))
```


#### Where is the disparity occuring?

```{r}


bin_diff <- state_bin_top  %>%
    group_by(state) |> 
    select(state,measure_name,order) |>
    pivot_wider(names_from=measure_name,values_from=order) |>
    mutate(top_rank = top_rank_us - top_rank_foreign)
    
plot_usmap( data = bin_diff, values = "top_rank", color = "white" ) + 
  scale_fill_continuous(
    low = "white", high = "blue", name = "Differential between US born versus foreign born wages", label = scales::comma
  ) + 
  labs(title = "Bin Difference (US - Foreign)",subtitle='Differential Strength (positive values:US>Foreign; negative: Foreign>US)') +
  theme(legend.position = "right")

```
