---
title: "Assignment 1"
subtitle: "Benjamin Inbar"
output: html_document
date: "August 31, 2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Pulling in Covid approval polling data.

This [data](https://github.com/fivethirtyeight/covid-19-polls) piqued my interest as it collated poll data and allows us to directly compare Trump vs. Biden's handling of the Coronavirus pandemic.

First, let's read in the data from Github, and check that all is in order, since this is my first time using R!
```{r}
data <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/covid-19-polls/master/covid_approval_polls.csv")
class(data)
head(data)
```

</div><div style="margin-top:50px;">
#### Great. Let's check column names only.
This might give us an idea of which we want to select for further transformation.
```{r}
colnames(data)
```

</div><div style="margin-top:50px;">
#### Let's say we want to run some predictive analytics. 
Let's subset or grab relevant columns to that end. Note that _**subject**_ would be our target or independent variable.

```{r}
trimmed <- subset(data, select=c(pollster, sample_size, population, party, subject, approve, disapprove))
head(trimmed)
```

</div><div style="margin-top:50px;">
#### A small transformation.
Make data values easier to understand for the 'population' variable.

```{r}
trimmed['population'][trimmed['population'] == 'a'] <- 'adult'
trimmed['population'][trimmed['population'] == 'rv'] <- 'registered voter'
trimmed['population'][trimmed['population'] == 'lv'] <- 'likely voter'
trimmed[1:20, 'population']
```

</div><div style="margin-top:50px;">
#### Future directions.
This was fun, but there are plenty of ways we could delve in, particularly if we are trying to build a predictive model. The first thing I'd do is run a correlation matrix on the selected variables.