---
title: "Assignment 4"
subtitle: "Ben Inbar"
output: html_document
date: "October 3, 2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="margin-top:50px;">
#### Read-in, pivot, and analyze flight delays!

<div style="margin-top:50px;">
First, read in the .csv and check.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
data <- read.csv("https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%204/Arrival%20delays.csv")
data
```

<div style="margin-top:50px;">
Clean. Drop NA rows, and replace blank values in Airline column with NA and fill them (since did not default these to NA on read-in).
```{r}
colnames(data)[1] <- "Airline"
colnames(data)[2] <- "Disposition"
data <- drop_na(data)
data[data==''] <- NA
data <- fill(data, Airline)
data
```

<div style="margin-top:50px;">
Pivot to a longer data format, and then, crucially, chain a pivot to wider format for the "on time" and "delayed" row values to that. Thanks to classmates for inspiration on the chaining of the pivot_wide.
```{r}
longified <- pivot_longer(data, cols=c(3:7), names_to="Destination", values_to="Number of flights") |> pivot_wider(names_from="Disposition", values_from="Number of flights")
longified
```

<div style="margin-top:50px;">
Find percent delayed, mutate to dataframe, and graph grouped by destination and binned by airline to get an overview of most delayed flights by destination and airline.
San Francisco, despite having fewer flights than Seattle, has the most delays!
```{r}
longified <- mutate(longified, `Percent delayed` = 100*(delayed / `on time`))
ggplot(data=longified, aes(x=Destination, y=`Percent delayed`, fill=Airline)) +
    geom_bar(stat="identity", position="dodge") +
    scale_fill_brewer(palette="Paired") +
    ggtitle("Flights") +
    theme(axis.title.y=element_blank()) +
    coord_flip()
```

<div style="margin-top:50px;">
Take a quick look at means of percent delays. Looks like AM West could use with some streamlining to reduce delays!
```{r}
longified |> group_by(Airline) |> summarize(`Mean percent delayed`=mean(`Percent delayed`))
```
