---
title: "Assignment 5"
author: "Benjamin Inbar"
date: "October 13, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### Write R code, using your packages of choice, to load the information from each of the three sources into separate R data frames.

```{r warnings=FALSE, messages=FALSE}
library(rvest)
library(jsonlite)
library(XML)
library(RCurl)
library(knitr)
```

<div style="margin-top:50px;">
<br>
Pull from <a href="https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%205/Books.html">HTML</a>. Use the rvest package, and convert the html table from the html read-in to a dataframe.
```{r}
html <- read_html("https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%205/Books.html")
data <- html |> html_table()
html_df <- as.data.frame(data)
kable(html_df)
```

<div style="margin-top:50px;">
<br>
Pull from <a href="https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%205/Books.json">JSON</a>. Use the jsonlite package and set 'simplifyVector=TRUE' to read directly into a dataframe.
```{r}
json <- read_json("https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%205/Books.json", simplifyVector=TRUE)
kable(json)
```

<div style="margin-top:50px;">
<br>
Pull from <a href="https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%205/Books.xml">XML</a>. Use the RCurl library to pull the page, which then allows the xml parser to work.
```{r}
url <- getURL("https://raw.githubusercontent.com/beninbar/DATA-607/main/Assignment%205/Books.xml")
xml <- xmlToDataFrame(url)
kable(xml)
```
