---
title: "Project 1"
subtitle: "Ben Inbar"
output: html_document
date: "2022-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players: Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents.

<div style="margin-top:50px;">
First, let's import chess player .txt file (read in as delimited file) and explore what we have. 195 rows/observations. 1 massive column.
```{r warning=FALSE, message=FALSE}
chessdata <- read.delim(file = "https://raw.githubusercontent.com/beninbar/DATA-607/main/Project%201/7645617.txt", header = TRUE)
class(chessdata)
library(knitr)
head(chessdata)
chessdata[1,]
chessdata[2,]
chessdata[3,]
chessdata[4,]
chessdata[5,]
chessdata[6,]
chessdata[7,]
```

<div style="margin-top:50px;">
Drop "header" rows to make Regex slightly easier.
```{r}
chessdata = chessdata[-c(1,2),]
head(chessdata)
```

<div style="margin-top:50px;">
Get players using RegEx, and remove blank returns from resulting matrix and coerce to vector.
```{r warning=FALSE, message=FALSE}
library(tidyverse)
players <- str_extract_all(chessdata, "[:upper:]+\\s[:upper:]*\\s?[:upper:]+.?\\s?[:upper:]+", simplify=TRUE)
players <- players[!apply(players == "", 1, all), ]
players
```

<div style="margin-top:50px;">
Get players' state.
```{r}
state <- str_extract_all(chessdata, "\\s\\s\\b[:upper:]{2}\\b", simplify=TRUE)
state <- state[!apply(state == "", 1, all), ]
state <- sub('  ', '', state)
state
```

<div style="margin-top:50px;">
Get total points.
```{r}
points <- str_extract_all(chessdata, "[:digit:]\\.[:digit:]", simplify=TRUE)
points <- points[!apply(points == "", 1, all), ]
points
```

<div style="margin-top:50px;">
Get player pre-ratings. Convert character vector to numeric to work with downstream average rating.
```{r}
pre_rating <- str_extract_all(chessdata, "(?<=R:\\s{1,2})[:digit:]{3,4}(?=P*)", simplify=TRUE)
pre_rating <- pre_rating[!apply(pre_rating == "", 1, all), ]
pre_rating <- as.numeric(pre_rating)
pre_rating
```

<div style="margin-top:50px;">
To get the average pre-chess rating of opponents, first get opponent IDs.
```{r}
opponent_IDs <- str_extract_all(chessdata, "(?<=\\s{1,2})[:digit:]{1,2}\\|", simplify=TRUE)
opponent_IDs <- opponent_IDs[!apply(opponent_IDs == "", 1, all), ]
opponent_IDs <- sub('\\|', '', opponent_IDs)
head(opponent_IDs)
```

<div style="margin-top:50px;">
Then, convert opponent_IDs to numeric matrix, and substitute these opponent_IDs for pre-chess rating of each ID matching to the pre_rating vector above. Finally, get row means for the average pre-chess rating.
```{r}
rating <- matrix(as.numeric(opponent_IDs), ncol=ncol(opponent_IDs))
head(rating)
rating <- matrix(pre_rating[rating], ncol=ncol(rating))
head(rating)
avg_rating_opponents <- rowMeans(rating, na.rm = TRUE, dims = 1)
avg_rating_opponents
```

<div style="margin-top:50px;">
Finally, stitch together into a dataframe, and export as a CSV. This [CSV can be found at my Github.](https://github.com/beninbar/DATA-607/blob/main/Project%201/chessdata_csv.csv)
```{r}
chessframe <- data.frame(players, state, points, pre_rating, avg_rating_opponents)
head(chessframe)
write.csv(chessframe, "C:/Users/bpinb/OneDrive/Desktop/Coursework/DATA 607 Data Acquisition and Management/Project 1/chessdata_csv.csv", row.names = FALSE)
```
