---
title: "Assignment 6_NYT API"
author: "Benjamin Inbar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<div style="margin-top:50px;">
##### Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

<div style="margin-top:50px;">
Call in httr and jsonlite libraries.
```{r warnings=FALSE, messages=FALSE}
library(httr)
library(jsonlite)
library(knitr)
```

<div style="margin-top:50px;">
Hide my API key by calling it in from a file located in the directory related to this project.
```{r warning=FALSE, messages=FALSE}
setwd("C:/Users/bpinb/OneDrive/Desktop/Coursework/DATA 607 Data Acquisition and Management/Assignment 6_NYT_API")
key <- read.delim("NYT_API.txt")
key <- colnames(key)
```

<div style="margin-top:50px;">
Get NYT movie reviews of Titanic.
```{r}
url <- paste0("https://api.nytimes.com/svc/movies/v2/reviews/search.json?query=titanic&api-key=", key)
NYT <- GET(url)
NYT
```

<div style="margin-top:50px;">
Extract raw NYT data to text, then to JSON, then coerce to a dataframe.
```{r warning=FALSE, messages=FALSE}
text <- content(NYT, "text")
text_as_json <- fromJSON(text)
titanic_df <- as.data.frame(text_as_json)

kable(titanic_df)
```
