---
title: "Extra credit 2"
subtitle: "Benjamin Inbar"
output: html_document
date: "2022-09-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

::: {style="margin-top:50px;"}
#### Welcome to my manually configured movie recommender system, based on movie reviews from 6 friends. This draws from a MySQL database.
:::

::: {style="margin-top:50px;"}
###### Access SQL movie reviews from previous (slightly modified). Connect to the SQL database using the "blinded user" created in the SQL script, and check connection was successful.

```{r warning=FALSE, message=FALSE}
library(RMySQL)
movies = dbConnect(MySQL(), user='ACatlin', password='BenAssignment2', dbname='movie_reviews')
dbListFields(movies, 'Results')
```
:::

::: {style="margin-top:50px;"}
###### Grab the data from the Results table and place it into a dataframe. GetQuery not SendQuery.

```{r}
get_table_data_df = dbGetQuery(movies, "select * from Results")
class(get_table_data_df)
head(get_table_data_df)
```
:::

::: {style="margin-top:50px;"}
###### Flatten data frame, calculate means across rows and columns, as well as the overall mean, bind those to a new flattened matrix array. Use Knitr's 'kable' library to prettify.

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(knitr)
flattened <- xtabs(Score ~ Name + Movie, get_table_data_df, addNA = TRUE)
rowmeans <- rowMeans(flattened, na.rm = TRUE, dims = 1)
colmeans <- colMeans(flattened, na.rm = TRUE, dims = 1)
withmeans <- cbind(flattened, rowmeans)
withmeans <- rbind(withmeans, colmeans)
overallmean <- colMeans(get_table_data_df['Score'], na.rm = TRUE, dims = 1)
withmeans[7, 7] = overallmean
kable(withmeans)
```
:::

::: {style="margin-top:50px;"}
###### Calculate the differences in user means and movie means, and bind those to (another copy of) the matrix.

```{r warning=FALSE, message=FALSE}
useravg_againstall <- rowmeans - overallmean
movieavg_againstall <- colmeans - overallmean
final_with_NAs <- cbind(withmeans, useravg_againstall)
final_with_NAs <- rbind(final_with_NAs, movieavg_againstall)
kable(final_with_NAs)
```
:::

::: {style="margin-top:50px;"}
###### Calculate rating for unseen item using difference variables calculated and impute to that index and column location in the matrix.

Several important steps here: 1. Make copy for comparison to NAs table and final table with imputed values 2. Get nulls 3. For nulls in table "final", impute mean + movie average differential + user average differential for that specific row/column index. 4. Check final. 5. Print just null values.

```{r}
final <- final_with_NAs
nulls <- which(is.na(final), arr.ind=TRUE)
final[nulls] <- overallmean + movieavg_againstall[nulls[,2]] + useravg_againstall[nulls[,1]]
kable(final)
final[nulls]
```
:::

<div style="margin-top:50px;">
### Based on these values, I would pick the highest baseline estimate of 4.226667 and recommend that 'Jacqueline' see 'Nope.'
